# STAR Example Intent-to-treat Report

```{r opts, echo=FALSE, results='hide'}
opts_chunk$set(echo=FALSE, warnings=FALSE, message=FALSE, comment = "")
opts_knit$set(root.dir='../../../')
```


```{r Rdata ,results='hide'}
# source('data/starRmake.R')
source('analyses/prePost.R')
source('analyses/stdPrePost.R')
library(reporttools)
library(psych)
library(Gmisc)
library(texreg)
library(tables)
# create data frames for descriptive statistics tables
factor.df <- star[ ,c('gender', 'ethnicity', 'schoolk','birth', 'lunchk')]
numeric.df <- star[ ,c('readk', 'read1', 'mathk', 'math1')]
```

## Introduction

This is an example report of the _STAR_ data. The data consists of `r nrow(star)` cases and `r ncol(star)` variables. This example is a simplified version of the types of internal lab reports that can be generated using R and the `knitr` package to produce reproducible analyses. 


## Descriptive Statistics
### Table 1.
```{r crosstab1, results='asis'}
Mean <- function(x) mean(x, na.rm=TRUE)
SD <- function(x) sd(x, na.rm=TRUE)
Min <- function(x) min(x, na.rm=TRUE)
Max <- function(x) max(x, na.rm=TRUE)

html(tabular((mathk + readk)*(Mean + SD) ~ schoolk + 1, star))

```

### Table 2.
```{r tab2, results='asis', warning=FALSE}


html(tabular(All(star) ~ (n=1) + Mean + SD + Min + Max , star))

```
### Descriptive table
```{r descriptiveNumeric, results='asis'}
dat <- round(as.data.frame((psych::describe(star))),2)
htmlTable(dat) # Gmisc package
```

## Barplots

```{r barplot gender, results='asis'}
barplot(table(star$gender), main='Gender')
barplot(table(star$ethnicity), main='Ethnicity')
barplot(table(star$birth), main='Birth')
barplot(table(star$stark), main='treatment condition')

```

## Histograms

```{r histograms, results='asis'}
hist(star$readk, col='skyblue', main='K Reading')
hist(star$mathk, col='skyblue', main='K Math')
hist(star$read1, col='skyblue', main='1st Reading')
hist(star$math1, col='skyblue', main='1st Math')
```


```{r pairsplot, results='asis', fig.pos='h', fig.width=12, fig.height=12}
pairs.panels(star[ ,-1], hist.col='steelblue')
```

## Intent-to-treat Analyses
### Pretest-posttest Regressions
```{rprePost, results='asis'}
htmlreg(list(readmod.pp, mathmod.pp), custom.model.names=c('Reading', 'Math'),
       float.pos='h', caption='Unstandardized ITT Models')
```

```{r stdPrePost, results='asis'}
htmlreg(list(readmod.spp, mathmod.spp), custom.model.names=c('Reading', 'Math'),
       float.pos='h', caption='Standardized ITT Models')
```


## R Session Information
```{r sessionInfo, results='markup'}
sessionInfo()
```




