Basic Data Transformations in R
========================================================
## Setup
### Options chunk
```{r opts}
opts_knit$set(root.dir='../../../')
```

### Data chunk
```{r}
# source('data/starRmake.R')
star <- read.csv(file='data/star.csv', header=TRUE)
```

## Transforming the Variable Class
Data transformations focus on taking raw variables and modifying them so they are more appropriate for the planned analysis.  We hav previously discussed how to explore your data (e.g. evaluate distributions, look for outliers). In this tutorial we consider functions useful in making changes to variables once they are identified as problematic.

### Factor to Numeric
```{r}
class(star$schoolk)
barplot(table(star$schoolk))
star$schoolk.n <- as.numeric(star$schoolk)
hist(star$schoolk.n, col='skyblue')
```

### Numeric to Factor

```{r}
# look at the readk variable (numeric)
hist(star$readk, col='skyblue')
# set break points
brks <- c(0, 425, 475, 525, Inf)
star$readk.fac <- cut(star$readk, breaks=brks)
class(star$readk.fac)
barplot(table(star$readk.fac))
# with labels
star$readk.fac <- cut(star$readk, breaks=brks, 
                      labels=c('low', 'low med','hi med', 'hi' ))

barplot(table(star$readk.fac))
```

## Recoding  Variables
```{r }
library(car)
# factor
star$readk.fac <- recode(star$readk.fac, "'low med'='med'; c('hi med', 'hi')= 'hi'")
barplot(table(star$readk.fac))
# reorder factors
star$readk.fac <- factor(star$readk.fac, levels=c('low', 'med', 'hi'))
# or, using original recode function with levels argument
star$readk.fac <- recode(star$readk.fac, "'low med'='med'; c('hi med', 'hi')= 'hi'", levels=c('low', 'med','hi'))
barplot(table(star$readk.fac))
# another factor
barplot(table(star$schoolk))
star$schoolk.3 <- recode(star$schoolk, "c('suburban', 'urban')='urban'")
barplot(table(star$schoolk.3))

# numeric 
star$schoolk.3n <- recode(star$schoolk.n, "3:4=4")
barplot(table(star$schoolk.3n))
```

## Dummy Code variables
```{r}
# YOU DON'T NEED TO FOR REGRESSION IN R!!
table(star$stark)
mod <- lm(read1 ~ stark, star)
summary(mod)

# Dummy coding with the dummies package
# install.packages('dummies')
library(dummies)
stark.dumdf <- dummy(star$stark)
star <- dummy.data.frame(star, names='stark')
```

## Using  ```dplyr``` Package
```{r}
library(dplyr)
head(star)

df <- tbl_df(star)

fem.df <- filter(df, gender=='female')
fem2.df <- df[df$gender=='female', ]
male.df <- filter(df, gender=='male')

fem.df <- arrange(fem.df, readk)
head(fem.df)
```
